Grounding the interaction: anchoring situated discourse in everyday human-robot interaction

### 2011-07-15 paper_knowledge.tex
Raquel changes to the experiments and conclusion parts

Third, the same symbolic knowledge base enables richer language capabilities
for the robot.  We propose a new approach to natural language grounding that is
robust, situated and more generic than what can be found in previous work
on situated language grounding. We present several examples that include a
generic recognition and semantic validation of thematic roles (\ie, the
semantic roles played by the different nominal groups around a verb), which
can be translated into input for a symbolic task planner, or disambiguation
based on attention foci.

@@

Section 5 presents three use-cases that were conducted on three different
robotic platforms: the \emph{naming} experiment, where a robot anchors new
knowledge in its model through verbal interaction; the \emph{Spy Game} where
either the user or the robot tries to guess which object the other player is
thinking of; and a more complex experiment situated in an everyday setup, where
the robot builds models for several agents and interacts with the users using
this knowledge.

@@

Anchoring perceptions in a symbolic model requires first perception abilities
and second, a symbolic assessment of the percepts. In this section we present
SPARK (\emph{SPAtial Reasoning \& Knowledge}), a situation assessment reasoner
that generates relevant symbolic information from the geometry of the
environment with respect to relations between objects and human capabilities.
Moreover, the notion of \emph{Perspective
Taking}~\cite{Flavell1992,Tversky1999} is employed in the heart of the
reasoning to provide the robot with the ability to put itself at the human's
place and to reason about the world from different perspectives.

As mentioned in the introduction, this paper do not focus on the sensor-level
perception. We rather assume that the perception of the humans and the objects
is provided as a list of unique identifiers with associated 3D meshes and 6DOF
poses.

@@

There are a number of common properties for a robot and a human related to
their capabilities in a given situation: they can both reach, grasp, look,
point, etc. In our context, we group robots and humans into a single category.
Thus, we define agents as entities that can act in the environment and
manipulate it. In this work we focus on the following capabilities from each
agent's perspective\footnote{Note that each of the capabilities described are
computed from each agent point of view, and therefore, also stored in different
models in ORO for further use at the decisional level.}.

@@

Experiments have been conducted on two different platforms: the
\emph{Rosie} manipulator from the Technical Universtiy of Munich, a dual-arm,
holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot from
LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
Genom/pocolibs} stack.

@@

\begin{figure}
\centering

\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/kimp1.jpg}} \hspace{0.2em}
\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/rviz.png}}
\flushleft
\hspace{1.8cm} (a) \hspace{3.9cm} (b)
\caption{(a) Rosie, looking for objects it may know, and (b) viewed in Willow
Garage's RViz. The clusters of point are given a unique identifier by the
perception that allows the supervision to create the link between the physical
objects and their symbolic representation in ORO.} 
\label{fig|kimpwatching}
\end{figure}

@@

The robot picks an unknown object from the table, shows it to the user, and
asks about its name and type (Figure~\ref{fig|kimpwatching}). The user
interacts with the robot through a dedicated bridge, allowing to chat with
the robot with a standard Jabber instant messaging client.
Figure~\ref{tab|transcript_kimp} shows a chat session with Rosie performing the
naming experiment. The human describes the object until the robot recognizes a
type it already knows: by asking in loop the human for the categories of an
object until it can connect it to a concept it already knows, the robot
accurately anchors perception in its symbolic model and it is able to reason
about it.  The learning process starts over again with each unknown object. To
further exploit the knowledge reasoner in this experiment, we ask the robot
which object do not belong to a typical breakfast table, \ie we ask for the odd
objects. In this case, it corresponds to objects that are neither of type
\concept{Tableware} nor \concept{Food}.

The supervision (\textsc{cram}\footnote{\textsc{cram} (Cognitive Robotic
Abstract Machine) is a RPL-derived framework for rapid development of cognitive
robot control programs we currently develop.}~\cite{Beetz2010}) automatically
updates the ORO server whenever an object enters or leaves the field of view.
%(the perception is based on the \textsc{CoP} framework~\cite{Klank2009}). 
Therefore, the integration of ORO can be seen as an extension to
the robot's belief state that not only contains abstract identifiers
of the internal object representation used in plans, but also the
semantics and roles of objects in the scenario.

@@

An unexpected example of what the symbolic reasoning layer brings to more
traditional robotic architectures emerged during the \emph{Naming} experiment:
the perception routines provided segmented blobs corresponding to objects along
with their colors to update the information in ORO. At some point, ORO suddenly
refused to add an object. What seemed at first a communication bug between
modules, was actually the consequence of a consistency check by ORO: because of
bad light conditions, the color recognition was not very reliable, and the same
object was set to have two different colors at the same time. Since the
\concept{hasColor} predicate we had decided to use was a functional predicate
(\ie objects can have only one color), ORO inferred an inconsistency and
discarded the new fact. This kind of logical failure can be used to improve
low-level perception results by ``closing the loop'' with high-level, symbolic
knowledge.

\subsection{``Spy Game'' experiment}

@@

\begin{figure}
\centering

\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/spy-game-real.jpg}} \hspace{0.2em}
\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/spy-game-mhp.jpg}}
\flushleft
\hspace{1.8cm} (a) \hspace{3.9cm} (b)
\caption{Spy game scenario: (a) Real environment and (b) 3D environment model, viewed in \textsc{SPARK}.}
\label{fig|spyGameScenario}
\end{figure}

@@

We have integrated the game in the LAAS-CNRS Jido robot~\cite{Alami1998a}.
Objects are identified through a tag-based vision approach\footnote{ARToolKit:
\url{http://www.hitl.washington.edu/artoolkit/}} and motion capture is used for
human tracking. Their descriptions regarding categories (type of object) and
features (color, shape) are manually given in advance. Spatial relationships
(\concept{front}, \concept{back}, \concept{left}, etc, and \concept{in},
\concept{on} and \concept{next to}) and visibility (only visible objects for
both agents can be considered in the game) are automatically computed on-line
by \textsc{SPARK} (the geometric reasoner). Figure~\ref{fig|spyGameExample}
shows an example of a round game.

\begin{figure}
\centering
	\begin{tabular}{r|p{6cm}}
		\chatN{human} & It is a tableware.\\
		\chatN{jido} & \emph{[retrieves possible objects: coke-bottle, yellow-bottle, orange-bottle, cup-with-handle]} \\
	 			& \emph{[keeps visible objects: coke-bottle, yellow-bottle, cup-with-handle]}\\
				& \emph{[obtains discriminants: type, color.]}\\
				& Which type of object is: bottle or cup? \\
		\chatN{human} & Bottle.\\
		\chatN{jido} & \emph{[obtains possible objects: coke-bottle, yellow-bottle.]}\\
				& \emph{[obtains discriminants: color.]}\\
				& What color the object is: red or yellow?\\
		\chatN{human} & Red.\\
		\chatN{jido} & \emph{[obtains possible objects: coke-bottle.]}\\
				& The object is the Coke bottle!	
	\end{tabular}\\
	\caption{Example of the robot playing Spy game.}
	\label{fig|spyGameExample}
\end{figure}

@@

In order to illustrate some of different reasoning abilities of the robot,
based on the approach presented in this paper, we have designed the following
daily life situation. Tom and Jerry are moving to London, so they are packing
things in boxes. The scenario takes places in the living-room, where Jido (our
robot) is observing while they move things here and there. To assess the
reasoning abilities of the robot they ask Jido for information (entered through
keyboard). Ideally, the robot should also perform actions when required (\eg
hand an object when asking ``give me...''). However, since it is out of the
scope of this work, we do not include any motion from the robot's side. Similar
to the Spy Game scenario, perception of objects is done through a tag-based
system and humans are detected through motion capture. We next describe in
detail three situations where we can follow the internal robot's reasoning and
the interaction with the users.

@@

First Tom enters the room and places on the table some of the things to pack: a
toolbox and two videos. He leaves one of the videos (``Jido-E'') inside one of
the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
table. We next describe how the situation assessment takes place, \ie how the
ontology is updated with the information obtained from the geometric reasoner
SPARK.

The initial information in the ontology corresponds to:

\begin{center}
    \begin{tabular}{l}
      \stmt{table type Table}\\
      \stmt{cardBoardBox type Box}\\
      \stmt{toolbox type Box}\\
      \stmt{videoTape1 type VideoiTape}\\
      \stmt{videoTape1 label "Lord of the Robots"}\\
      \stmt{videoTape2 type VideoTape}\\
      \stmt{videoTape2 label "Jido-E"}\\
      \stmt{Tom type Human}\\
      \stmt{Jerry type Human}
    \end{tabular}
\end{center}
\vspace{0.10cm}

There is a cardboard box on the table:

@@

Tom enters carrying several objects (Figure~\ref{fig|sitAssessRoman}a).
He places a toolbox and a video tape on the table, and another video tape
inside the box next to Jido and the leaves. (Figure~\ref{fig|sitAssessRoman}b):

@@

Since only one object is available, the robot infers
that the human refers to it and can eventually execute the command, \ie give
it to the human. Alternatively, the robot could first verify with the human if
that was the object being referred to or not before proceeding to execute the
action. Table~\ref{table|ptbeliefs} lists the robot's beliefs about itself and
its human partner involved in this situation.

@@

However both are visible and the cognitive ambiguity resolution cannot be
applied. The only option is to ask Jerry which box he is referring to: ``Which
box, the toolbox or the cardboard box?'' Jerry could now simply answer the
question. Instead, he decides to point at it while indicating: ``This box''
(Figure~\ref{fig|pointing}b). The robot's perception identifies the
\concept{cardBoardBox} as being pointed at and looked at by the human and
updates the ontology with this new information using a rule available in the
commonsense ontology ({\tt pointsAt(?agt, ?obj) $\land$ looksAt(?agt, ?obj)
$\to$ focusesOn(?agt, ?obj)}) The \textsc{Dialogs} module is then able to merge
both sources of information, verbal (``this'') and gestural to distinguish the
box Jerry refers to.

\begin{center}
\begin{tabular}{l}
\stmt{Jerry pointsAt carboardBox}\\
\stmt{Jerry looksAt carboardBox}\\
$\to$ \stmt{Jerry focusesAt carboardBox}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox]}
\end{tabular}
\end{center}

@@

\begin{itemize}
\item{Knowledge is explicitly stored in one central and consistent repository
of facts, accessible by all modules,}
\item{Knowledge is represented in a strict formalism (RDF statements) and
with a clearly defined vocabulary (stated in the {\tt commonsense.oro.owl}
ontology),}
\item{The first two points enable both a loosely-coupled
architecture where modules can very easily be removed or replaced by other ones
as long as they share the same semantics (modules are defined by the knowledge
they produce),}
\item{and a \emph{symbolic} reactive, event-driven approach
to supervision. By managing events at the same level as
the reasoner, we take full advantage of the inference abilities of ORO to
trigger events whose \textit{True} conditions can be inferred.}
\item{Finally, this architecture allows for merging very different knowledge
modalities in a single homogeneous environment, bringing mutual benefits to
components: for instance, the dialogue processing module can perfectly run
without any geometric perception, but its disambiguation routines can
transparently benefit from it when available (since richer symbolic
descriptions of objects are then available).}
\end{itemize}

Regarding the anchoring question, this architecture is
bidirectional: the components we described provide a \textit{bottom-up}
grounding process: SPARK and \textsc{Dialogs} constantly build and push new
symbolic contents about the world to ORO where it becomes accessible to
decisional layers. In parallel, ORO relies on reasoning in a \textit{top-bottom}
way to produce new facts that may trigger in return physical behaviours.

@@

The versatility of this paradigm could be illustrated by a simple imaginary
scenario with a blind robot and a deaf robot. The blind robot does not see (no
cameras or alike), but someone can verbally describe a scene to it, while the
deaf robot has a good vision system but can not hear anything.  Without any
changes to the software architecture above the language processing module or
the vision module, supervisors of both robot would be able to perform equally
well.

@@

We hope as well that our contribution may contribute to fill the gap between
robotics and psychology. Our architecture provides easier-to-reach entry points
to implement some classical psychology tests to robots. We presented
experiments focused on issues related to perspective taking. By explicitly
enabling independent modeling of the beliefs of each agent, our architecture is
especially well suited to set up \emph{False-Belief}
experiments~\cite{Leslie2000}, and we plan to further explore those kinds of
applications.

@@

For us, \emph{Grounding} is on the contrary a \emph{top-bottom} activity: the
robot needs to automatically bind a representation (for instance, a word
uttered by a human, an image taken from a camera, a sentence extracted from
Wikipedia) to an unambiguous, context-dependent, internal concept. This very
concept can (or not) be \textit{a priori} available to the robot as a
pre-loaded ontology (what we previously called the cultural background of the
robot).

@@

The three experiments that were presented in the paper all illustrate how the
robot make use of its embodied nature to establish a meaningful communication
with a human: because the robot and the human share the same physical
environment and they perceive each other, we are able to create a mutual
context.

Sloman, in~\cite{Sloman2009}, argues however that the strong focus on
embodiment in the robotics community has hindered progress towards natural
human-robot interaction. Our approach has hopefully made clear that, similar to
Beetz et al.~\cite{Beetz2010}, we do not consider embodiment \emph{per se}, outside of
a broader symbolic system: our architecture is not bound to the morphology or
the low-level sensori-motor capabilities of a specific agent.

However, we can build a model of the ``human point of view'' because the robot
sees the human, and is able to deduce, at least partially, what the human
perceives or not. This relies on the embodied nature of the interaction. We
infer that a human focuses on some object because it points at it, looks at it,
and besides, the object is visible to him. This relies as well on the embodied
nature of the interaction. And this allow us, in turn, to understand the
meaning of sentences like ``Give me this''.

@@

Let conclude this article by a short glance back at Roy's list of challenges
for human-robot dialogue: 
\begin{itemize}

	\item While more modalities (especially, deictic gestures and social gazes)
	can be added, we have proposed in this article a \emph{cross-modal
	representation system},

>>>>

Third, the same symbolic knowledge base enables richer language capabilities
for the robot.  We propose a new approach to natural language grounding that is
robust, situated and more generic than what can be found in previous work
on situated language grounding. We present several examples that include a
generic recognition and semantic validation of thematic roles (\ie the
semantic roles played by the different nominal groups around a verb), which
can be translated into input for a symbolic task planner, or disambiguation
based on attention foci.

@@

Section 5 presents three use-cases that were conducted on two different
robotic platforms: the \emph{Naming} experiment, where a robot anchors new
knowledge in its model through verbal interaction; the \emph{Spy Game} where
either the user or the robot tries to guess which object the other player is
thinking of; and a more complex experiment situated in an everyday setup, where
the robot builds models for several agents and interacts with the users using
this knowledge.

@@

Anchoring perceptions in a symbolic model requires perception abilities
and their symbolic assessments. In this section we present
SPARK (\emph{SPAtial Reasoning \& Knowledge}, \cite{Sisbot2011}), a situation assessment reasoner
that generates relevant symbolic information from the geometry of the
environment with respect to relations between objects, robots and humans.
Moreover, the notion of \emph{Perspective
Taking}~\cite{Flavell1992,Tversky1999} is employed at the heart of the
reasoner to provide the robot with the ability to put itself at the human's
place and to reason about the world from different perspectives.

As mentioned in the introduction, this paper does not focus on the sensor-level
perception. We rather assume that the perception of the humans and the objects
is provided as a list of unique identifiers with associated 3D meshes and 6DOF
poses.

@@

There are a number of common properties for a robot and a human related to
their capabilities in a given situation: they can both reach, grasp, look at,
point at, etc. In our context, we group robots and humans into a single category.
Thus, we define agents as entities that can act in the environment and
manipulate it. In this work we focus on the following capabilities from each
agent's perspective\footnote{Note that each of the capabilities described are
computed from each agent point of view, and therefore, also stored in different
models in ORO for further use at the decisional level.}:

@@

Experiments have been conducted on two different platforms: the
\emph{Rosie} manipulator from the Technical Universtiy of Munich, a dual-arm,
holonomic service robot, running the ROS\footnote{\emph{Robotic Operating
System}, \url{http://www.ros.org}} middleware, and the \emph{Jido} robot~\cite{Alami1998a} from
LAAS-CNRS, a similar, single-arm, service robot, running the LAAS's {\sc
Genom/pocolibs} stack.

@@

\begin{figure}
\centering
\includegraphics[width=0.45\columnwidth]{images/kimp1.jpg}
\includegraphics[width=0.45\columnwidth]{images/rviz.png}
%\parbox[c]{4.3cm}{\includegraphics[width=0.5\columnwidth]{images/kimp1.jpg}}
%\parbox[c]{4.1cm}{\includegraphics[width=4cm]{images/rviz.png}}
\flushleft
\hspace{1.8cm} (a) \hspace{3.9cm} (b)
\caption{(a) Rosie, looking for objects it may know, and (b) view from Willow
Garage's RViz. The clusters of point are given a unique identifier by the
perception that allows the supervision to create the link between the physical
objects and their symbolic representation in ORO.} 
\label{fig|kimpwatching}
\end{figure}

@@

The robot selects an unknown object from the table, shows it to the user, and
asks about its name and type (Figure~\ref{fig|kimpwatching}). The user
interacts with the robot through a dedicated bridge, allowing to chat with
the robot with a standard Jabber instant messaging client.
Figure~\ref{tab|transcript_kimp} shows a chat session with Rosie performing the
naming experiment. The human describes the object until the robot recognizes a
type it already knows. More precisely, the robot recursively asks the human for the categories the object belongs to until reaching a known one. At this point, the robot is able to connect the new object information to already known concepts. Therefore, the robot
accurately anchors perception in its symbolic model and it is able to reason
about it.  The learning process starts over again with each unknown object. To
further exploit the knowledge reasoner in this experiment, we ask the robot
which object do not belong to a typical breakfast table, \ie we ask for the odd
objects. In this case, it corresponds to objects that are neither of type
\concept{Tableware} nor \concept{Food}.

The supervision module (\textsc{cram}\footnote{\textsc{cram} (Cognitive Robotic
Abstract Machine) is a RPL-derived framework for rapid development of cognitive
robot control programs we currently develop.}~\cite{Beetz2010}) automatically
updates the ORO server whenever an object enters or leaves the field of view.
%(the perception is based on the \textsc{CoP} framework~\cite{Klank2009}). 
Therefore, the integration of ORO can be seen as an extension of
the robot's belief state that not only contains abstract identifiers
of the internal object representation used in plans, but also the
semantics and roles of objects in the scenario.

@@

An unexpected example of what the symbolic reasoning layer brings to more
traditional robotic architectures emerged during the \emph{Naming} experiment:
the perception routines provided segmented blobs corresponding to objects along
with their colors to update the information in ORO. At some point, ORO suddenly
refused to add an object. What seemed at first a communication bug between
modules, was actually the consequence of a consistency check by ORO: because of
bad light conditions, the color recognition was not very reliable, and the same
object was set to have two different colors at the same time. Since the
\concept{hasColor} predicate we had decided to use was a functional predicate
(\ie objects can have only one color), ORO inferred an inconsistency and
discarded the new fact. This kind of logical failure can be used to improve
low-level perception outcomes by ``closing the loop'' with high-level, symbolic
knowledge.

\subsection{\emph{Spy Game} experiment}

@@

\begin{figure}
\centering
\includegraphics[width=0.45\columnwidth]{images/spy-game-real.jpg}
\includegraphics[width=0.45\columnwidth]{images/spy-game-mhp.jpg}
%\parbox[c]{4.3cm}{\includegraphics[width=4.3cm]{images/spy-game-real.jpg}} \hspace{0.2em}
%\parbox[c]{4.1cm}{\includegraphics[width=4.1cm]{images/spy-game-mhp.jpg}}
\flushleft
\hspace{1.8cm} (a) \hspace{3.9cm} (b)
\caption{Spy game scenario: (a) Real environment and (b) 3D environment model, viewed in \textsc{SPARK}.}
\label{fig|spyGameScenario}
\end{figure}

@@

We have integrated the game on the LAAS-CNRS Jido robot.
Objects are identified through a tag-based vision approach\footnote{ARToolKit:
\url{http://www.hitl.washington.edu/artoolkit/}} and motion capture is used for
human tracking. Their descriptions regarding categories (type of object) and
features (color, shape) are manually given in advance. Spatial relationships
(\concept{front}, \concept{back}, \concept{left}, etc, and \concept{in},
\concept{on} and \concept{next to}) and visibility (only visible objects for
both agents can be considered in the game) are automatically computed on-line
by \textsc{SPARK} (the geometric reasoner). Figure~\ref{fig|spyGameExample}
shows an example of a round game.

\begin{figure}
\centering
	\begin{tabular}{r|p{6cm}}
		\chatN{human} & It is a tableware.\\
		\chatN{jido} & \emph{[retrieve possible objects: coke-bottle, yellow-bottle, orange-bottle, cup-with-handle]} \\
	 			& \emph{[keep visible objects: coke-bottle, yellow-bottle, cup-with-handle]}\\
				& \emph{[obtain discriminants: type, color.]}\\
				& Which type of object is: bottle or cup? \\
		\chatN{human} & Bottle.\\
		\chatN{jido} & \emph{[obtain possible objects: coke-bottle, yellow-bottle.]}\\
				& \emph{[obtain discriminants: color.]}\\
				& What color is the object: red or yellow?\\
		\chatN{human} & Red.\\
		\chatN{jido} & \emph{[obtains possible objects: coke-bottle.]}\\
				& The object is the Coke bottle!	
	\end{tabular}\\
	\caption{Example of the robot playing Spy game.}
	\label{fig|spyGameExample}
\end{figure}

@@

In order to illustrate some of different reasoning abilities of the robot presented in this paper, we have designed the following
daily life situation. Tom and Jerry are moving to London, so they are packing
things in boxes. The scenario takes places in the living-room, where Jido (our
robot) is observing while they move things here and there. To assess the
reasoning abilities of the robot they ask Jido for information (entered through
keyboard). Ideally, the robot should also perform actions when required (\eg
hand an object when asking ``give me...''). However, since it is out of the
scope of this work, we do not include any motion from the robot's side. Similar
to the \emph{Spy Game} scenario, perception of objects is done through a tag-based
system and humans are detected through motion capture. We next describe in
detail three situations where we can follow the internal robot's reasoning and
the interaction with the users.

@@

First Tom enters the room with some of the things they need to pack: a
toolbox and two videos. He leaves one of the videos (``Jido-E'') inside one of
the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
table. We next describe how the situation assessment takes place, \ie how the
ontology is updated with the information obtained from the geometric reasoner
SPARK.

The initial information in ORO corresponds to:

\begin{center}
    \begin{tabular}{l}
      \stmt{table type Table}\\
      \stmt{cardBoardBox type Box}\\
      \stmt{toolbox type Box}\\
      \stmt{videoTape1 type VideoTape}\\
      \stmt{videoTape1 label "Lord of the Robots"}\\
      \stmt{videoTape2 type VideoTape}\\
      \stmt{videoTape2 label "Jido-E"}\\
      \stmt{Tom type Human}\\
      \stmt{Jerry type Human}
    \end{tabular}
\end{center}
\vspace{0.10cm}

SPARK detects that there is a cardboard box on the table. It thus sends the fact to ORO:

@@

Tom enters carrying several objects (Figure~\ref{fig|sitAssessRoman}a).
He places a toolbox and a video tape on the table, and another video tape
inside the box next to Jido and the leaves. (Figure~\ref{fig|sitAssessRoman}b). The following facts are computed and sent to ORO:

@@

Since only one object is available now, the robot infers
that the human refers to it and can eventually execute the command, \ie give
it to the human. Alternatively, the robot could first verify with the human if
that was the object being referred to or not before proceeding to execute the
action. Table~\ref{table|ptbeliefs} lists the robot's beliefs about itself and
its human partner involved in this situation.

@@

However both are visible and the cognitive ambiguity resolution cannot be
applied. The only option is to ask Jerry which box he is referring to: ``Which
box, the toolbox or the cardboard box?'' Jerry could now simply answer the
question. Instead, he decides to point at it while indicating: ``This box''
(Figure~\ref{fig|pointing}b). The robot's perception identifies the {\tt
cardBoardBox} as being pointed at and looked at by the human and updates the
ontology with this new information using a rule available in the commonsense
ontology:
\begin{center}
\begin{tabular}{l}
\stmt{Jerry pointsAt carboardBox}\\
\stmt{Jerry looksAt carboardBox}\\
$\to$ \stmt{Jerry focusesAt carboardBox}\\
\end{tabular}
\end{center}
In the meantime, the \textsc{Dialogs} module is processing the human verbal input. When trying to resolve the reference ``this'' it is able to merge\footnote{Due to synchronization issues, the user should perform the gesture (pointing at) before answering the robot's question and maintain it until the resolution process takes place.} both sources of information, verbal and gestural, to distinguish the box Jerry refers to:
\begin{center}
\begin{tabular}{l}
\stmt{Jerry focusesAt ?obj}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox]}
\end{tabular}
\end{center}

@@

\begin{itemize}
\item{Knowledge is explicitly stored in one central and consistent repository
of facts, accessible by all modules,}
\item{Knowledge is represented in a strict formalism (RDF statements) and
with a clearly defined vocabulary (stated in the {\tt commonsense.oro.owl}
ontology),}
\item{The first two points enable both a loosely-coupled
architecture where modules can very easily be removed or replaced by other ones
as long as they share the same semantics (modules are defined by the knowledge
they produce),}
\item{and a \emph{symbolic} reactive, event-driven approach
to supervision. By managing events at the same level as
the reasoner, we take full advantage of the inference abilities of ORO to
trigger events whose \textit{True} conditions can be inferred.}
\item{Finally, this architecture allows for merging very different knowledge
modalities in a single homogeneous environment, bringing mutual benefits to
components. For instance, the dialogue processing module can perfectly run
without any geometric perception, but its disambiguation routines can
transparently benefit from it when available (since richer symbolic
descriptions of objects are then available).}
\end{itemize}

Regarding the anchoring question, this architecture is
bidirectional. The components we described provide a \textit{bottom-up}
grounding process: SPARK and \textsc{Dialogs} constantly build and push new
symbolic contents about the world to ORO where it becomes accessible to
decisional layers. In parallel, ORO relies on reasoning in a \textit{top-bottom}
way to produce new facts that may trigger in return physical behaviours.

@@

The versatility of this paradigm could be illustrated by a simple imaginary
scenario with a blind robot and a deaf robot. The blind robot does not see (no
cameras or alike), but someone can verbally describe a scene to it, while the
deaf robot has a good vision system but can not hear.  Without any
changes to the software architecture described, supervisors of both robots would be able to perform equally well.

@@

We hope as well that our contribution may contribute to fill the gap between
robotics and psychology. Our architecture provides easier-to-reach entry points
to implement some classical psychology tests to robots. We presented
experiments focused on issues related to perspective taking. By explicitly
enabling independent modeling of the beliefs of each agent, our architecture is
especially well suited to set up cognitive and psychological experiments (such as the \emph{False-Belief} experiment~\cite{Leslie2000}), which we plan to further explore.

@@

For us, \emph{grounding} is on the contrary a \emph{top-bottom} activity: the
robot needs to automatically bind a representation (for instance, a word
uttered by a human, an image taken from a camera, a sentence extracted from
Wikipedia) to an unambiguous, context-dependent, internal concept. This 
concept may (or may not) be \textit{a priori} available to the robot as a
pre-loaded ontology (what we previously called the cultural background of the
robot).

@@

The three experiments that were presented in the paper all illustrate how the
robot make use of its embodied nature to establish a meaningful communication
with a human. Mainly, because the robot and the human share the same physical
environment and they perceive each other, we are able to create a mutual
context.

Sloman, in~\cite{Sloman2009}, argues however that the strong focus on
embodiment in the robotics community has hindered progress towards natural
human-robot interaction. Our approach has hopefully made clear that, similar to
Beetz et al.~\cite{Beetz2010}, we do not consider embodiment \emph{per se} outside of
a broader symbolic system, \ie our architecture is not bound to the morphology or
the low-level sensori-motor capabilities of a specific agent.

However, we can build a model of the ``human point of view'' because the robot
perceives the human, and is able to deduce, at least partially, what the human
perceives or not. We infer that a human focuses on some object because it points at it, looks at it, and besides, the object is visible to him. This relies on the embodied
nature of the interaction. And this allows us, in turn, to understand the
meaning of sentences like ``Give me this''.

@@

Let us conclude this article by a short glance back at Roy's list of challenges
for human-robot dialogue: 
\begin{itemize}

	\item While more modalities (especially, deictic gestures and social gazes)
	can be added, we have actually proposed a \emph{cross-modal
	representation system},
